---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Libraries used
```{r echo=TRUE,message=FALSE}
library(dplyr)
library(lattice)
```

## Loading and preprocessing the data

```{r echo=TRUE}
data <- read.csv(unz("activity.zip","activity.csv"), na.strings="NA",
                 colClasses=c("integer","Date","integer"))

activity <- data %>% filter(!is.na(steps))
```

## What is mean total number of steps taken per day?

```{r echo=TRUE}
daily <- activity %>% group_by(date) %>% summarise(steps=sum(steps))
dmean   <- mean(daily$steps)
dmedian <- median(daily$steps)
hist(daily$steps, col="light blue", main="Daily Steps", xlab="steps")
abline(v=dmean,lty='dashed')
abline(v=dmedian,lty='dotted')
```

The mean total number of steps is: `r sprintf("%.4f",dmean)` and the median is `r dmedian`.

## What is the average daily activity pattern?

```{r echo=TRUE}
iactv <- activity %>% group_by(interval) %>% summarise(steps=mean(steps))
most_active_interval = iactv %>% filter(steps==max(steps))
with(iactv, plot(x=interval, y=steps, type="l"))
abline(v=most_active_interval[1,1],lti='dashed', col="red")
title(main="Daily activity pattern",sub="with most active interval")
```

The most active interval is: `r most_active_interval[1,1]` with a mean of `r  most_active_interval[1,2]` steps.

## Imputing missing values

There are a total of `r dim(data)[1]-dim(activity)[1]` missing values in the database.

To avoid bias in the calculations we're going to fill in the missing values using the mean of the missing data interval.

```{r echo=TRUE}
fill_in1 <- filter(data,is.na(steps)) %>% inner_join(iactv,by="interval") %>%
  select(steps=steps.y, date, interval) %>% rbind(activity)
```

```{r echo=TRUE}
daily1 <- fill_in1 %>% group_by(date) %>% summarise(steps=sum(steps))
dmean1   <- mean(daily1$steps)
dmedian1 <- median(daily1$steps)
hist(daily1$steps, col="light blue", main="Daily Steps", xlab="steps")
abline(v=dmean1,lty='dashed')
abline(v=dmedian1,lty='dotted')
```

The mean total number of steps is: `r sprintf("%.4f",dmean1)` and the median is `r sprintf("%.4f",dmedian1)`.

## Are there differences in activity patterns between weekdays and weekends?

Add weekdays as a factor, notice that I have used strftime to get the days as characters 0 to 6
where 0 = sunday, this is to avoid protability problems due to locale settings.

```{r echo=TRUE}
# Compute weekday factor
weekday <- sapply( strftime(fill_in1$date,"%w"), function(x) if (x==0 | x == 6) {"weekend"} else {"weekday"} )
wactiv <- mutate(fill_in1, weekday = as.factor( weekday )) %>% group_by(weekday,interval) %>% summarise(steps=mean(steps))
```

The following graph shows the activity difference between weekdays and weekends

```{r echo=TRUE}
xyplot(steps ~ interval | weekday, data = wactiv, layout=c(1,2), type="l")
```

